---
title: "Some Hospital Background"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Econ 771, Fall 2022
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{mathtools}
  - \usepackage{xspace}
  - \usepackage{pgfplots}
output:
#  html_document: default
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
#knit: pagedown::chrome_print
---

<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, magick,
               scales, plotly, gganimate, cobalt, MatchIt, ggthemes, here, latex2exp,
               hettreatreg, pdftools)
set.seed(12345)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


# Outline
1. [Admin Updates](#admin)

2. [Ownership Structure](#ownership)

3. [Payment and Coding](#payment)

4. [More Recent Interest](#recent)


---
# Quick notes

- Data for assignment 1 on our shared OneDrive folder, link is on Canvas
- Revisit coefficient versus ATE from prior class


<!-- New Section -->
---
class: inverse, center, middle
name: ownership

# Hospital Ownership Structure

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Ownership types

1. Private not-for-profit: About 60%

2. For-profit: About 20%

3. State and local gov't: About 20%

<br>
<div class="smalltext">Source: <a href="https://www.aha.org/statistics/fast-facts-us-hospitals">AHA Fast Facts</a></div>

---
# What does NFP mean?
From an economics perspective:
- Hospital assumed to maximize some objective function, $u(q,z)$, subject to a production constraint

- $q$ denotes quantity of care and $z$ denotes quality of care

- Production is constrained by the break-even condition

---
# What does NFP mean?
From a practical perspective:
- Profits must be re-invested into the hospital

- Must show "community benefit" (no consensus definition...includes uncompensated care, services to Medicaid, and certain specialized services that are generally unprofitable)

- No taxes! and tax-free bonds

---
# NFP Production Constraint
```{tikz echo=FALSE, fig.cap="Demand and Cost Changes from Increase in Quality", fix.ext='png'}
\usetikzlibrary{decorations,arrows,shapes,positioning}
\usetikzlibrary{calc,patterns}
\begin{tikzpicture}[scale=.6]
\draw [thick] (0,0) -- (8,0);
\draw [thick] (0,0) -- (0,5);
\node [right] at (8,0) {\tiny $x$};
\node [above] at (0,5) {\tiny \$};
\draw [thick] (1.5,4.6) -- (7.4,1.3);
\node [left] at (7.2,1.2) {\tiny $D_{H}$};
\draw [thick] (1,3.4) -- (5.3,1);
\node [left] at (5,1) {\tiny $D_{L}$};
\draw (.5,3.5) to [out=300,in=200] (6,2.5);
\node [below] at (.5,3.2) {\tiny $AC_{L}$};
\draw (.8,5) to [out=300,in=200] (7.2,3.5);
\node [below] at (7,3.4) {\tiny $AC_{H}$};
\draw[dashed](3.5,2)--(3.5,0);
\draw [fill] (3.5,2) circle [radius =0.05];
\draw[dashed](4.4,3)--(4.4,0);
\draw [fill] (4.4,3) circle [radius =0.05];
\end{tikzpicture}
```

---
# NFP Proeuction Constraint
```{tikz echo=FALSE, fig.cap="PPC from Demand and Cost Intersection", fix.ext='png'}
\usetikzlibrary{decorations,arrows,shapes,positioning}
\usetikzlibrary{calc,patterns}
\begin{tikzpicture}[scale=.5]
\draw [thick] (0,0) -- (8,0);
\draw [thick] (0,0) -- (0,7);
\node [right] at (8,0) {\tiny $x$};
\node [above] at (0,7) {\tiny \$};
\draw [thick] (2.5,5.6) -- (7.8,2.5);
\node [left] at (7.8,2.5) {\tiny $D_{HH}$};
\draw [thick] (1.5,4.6) -- (7.4,1.3);
\node [left] at (7.2,1.2) {\tiny $D_{H}$};
\draw [thick] (1,3.4) -- (5.3,1);
\node [left] at (5,1) {\tiny $D_{L}$};
\draw (.5,3.5) to [out=300,in=200] (6,2.5);
\node [below] at (.5,3.2) {\tiny $AC_{L}$};
\draw (.8,5) to [out=300,in=200] (7.2,3.5);
\node [above] at (7,3.6) {\tiny $AC_{H}$};
\draw (1.1,6) to [out=300,in=200] (7,5.5);
\node [above] at (7,5.6) {\tiny $AC_{HH}$};
\draw [fill] (3.5,2) circle [radius =0.05];
\draw [fill] (4.4,3) circle [radius =0.05];
\draw [fill] (4.05,4.75) circle [radius =0.05];
\draw [very thick] (1.8,1) .. controls (4,2) and (5.5,3.2) .. (3.5,5.5);
\end{tikzpicture}
```

---
# Non-profit hospitals and tax benefits
- <span>&#36;</span>24.6 billion in tax exemptions in 2011

- <span>&#36;</span>62.4 billion in "community benefits"<br>

- [Washington Post Article](https://www.washingtonpost.com/national/health-science/value-of-tax-breaks-for-nonprofit-hospitals-doubled-in-a-decade/2015/06/17/4162c640-1450-11e5-9ddc-e3353542100c_story.html?utm_term=.4f877a72a09f)<br>


---
# What is a non-profit hospital?
The real question is...what is the hospital's objective function?

- For-profit in disguise
- Output maximizers
- Social welfare maximizers
- Something else?


<!-- New Section -->
---
class: inverse, center, middle
name: payment

# Payments and Coding

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Organizing a Claim
Inpatient:
- DRG codes (e.g., DRG 470, Major Hip and Knee without MCC)
- Procedure codes (ICD-9 / ICD-10-PCS)
- Diagnosis codes (ICD-9 / ICD-10-CM)

---
# Organizing a Claim
Outpatient:
- APCs
- HCPCS codes
- Diagnosis codes (ICD-9 / ICD-10-CM)

---
# Organizing a Claim
Office:
- CPT/HCPCS codes
- Diagnosis codes (ICD-9 / ICD-10-CM)


<!-- New Section -->
---
class: inverse, center, middle
name: recent

# More Recent Interest

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# From today's NBER alerts

- [Private Provision of Public Services](https://www.nber.org/papers/w30390)
- [Governance of Non-Profits and their Social Impact](https://www.nber.org/papers/w30391)