---
title: "Instrumental Variables in the Modern Age"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Econ 771, Fall 2022
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{tikz}
  - \usepackage{pgf}
  - \usepackage{pgfarrows}
  - \usepackage{pgfnodes}
  - \usepackage{pgfautomata}
  - \usepackage{pgfheaps}
  - \usepackage[absolute,overlay]{textpos}  
  - \usepackage{mathtools}
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
bibliography: '../syllabus/BibTeX_Library.bib'      
---
class: inverse, center, middle



<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, fixest, readstata13,
               xaringanExtra, webshotm, ggthemes, gganimate, plotly, ivpack, ivreg)
set.seed(1234)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```

# Common IV Designs

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Judge Fixed Effects

- Many different possible decision makers
- Individuals randomly assigned to one decision maker
- Decision makers differ in leniency of assigning treatment
- Common in crime studies due to random assignment of judges to defendants

---
# Judge Fixed Effects

Aizer and Doyle (2015), QJE, "Juvenile Incarceration, Human Capital, and Future Crime: Evidence from Randomly Assigned Judges"
- Proposed instrument: propensity to convict by the judge
- Idea: judge has some fixed leniency, and random assignment into judges introduces exogenous variation in probability of conviction
- In practice: judge assignment isn't truly random, but it **is** plausibly exogenous

---
# Judge Fixed Effects

Constructing the instrument:
- Leave-one-out mean $$z_{j} = \frac{1}{n_{j} - 1} \sum_{k \neq i}^{n_{j}-1} JI_{k}$$
- This is the mean of incarceration rates for judge $j$ when excluding the current defendant, $i$
- Could also residualize $JI_{k}$ (remove effects due to day of week, month, etc.)

---
# Judge Fixed Effects

- Common design: possible in settings where some influential decision-makers exercise discretion and where individuals can't control the match
- Practical issue: Use jackknife IV (JIVE) to "fix" small-sample bias
    - JIVE more general than judge fixed effects design
    - Idea: estimate first-stage without observation $i$, use coefficients for predicted endogeneous variable for observation $i$, repeat
    - May improve finite-sample bias but also loses efficiency
- Biggest threat: monotonicity...judges may be more/less lenient in different situations or for different defendants

---
# Bartik (shift-share) Instruments

- Named after Timothy Bartik, traced back to Perlof (1957)
- Original idea: Estimate effect of employment growth rates on labor-market outcomes
    - Clear simultaneity problem
    - Seek IV to shift labor demand

---
# Bartik Instruments

Decompose observed growth rate into:
1. "Share" (what extra growth would have occurred if each industry in an area grew at their industry national average)
2. "Shift" (extra growth due to differential growth locally versus nationally)


---
# Bartik Instruments

- Want to estimate $$y_{l} = \alpha + \delta I_{l} + \beta w_{l} + \varepsilon_{l}$$ for location $l$ (possibly time $t$ as well)
- $I_{l}$ reflect immigration flows
- $w_{l}$ captures other observables and region/time fixed effects


---
# Bartik Instruments

Instrument:

$$B_{l} = \sum_{k=1}^{K} z_{l,k} \Delta_{k},$$

- $l$ denotes market location (e.g., Atlanta), country, etc. (wherever flows are coming *into*)
- $k$ reflects the source country (where flows are coming out)
- $z_{lk}$ denotes the **share** of immigrants from source $k$ in location $l$ (in a base period)
- $\Delta_{k}$ denotes the **shift** (i.e., change) from source country into the destination country as a whole (e.g., immigration into the U.S.)

---
# Other Examples

$$\begin{align}
B_{l} &= \sum_{k=1}^{K} z_{lk} \Delta_{k},\\
\Delta_{lk} &= \Delta_{k} + \tilde{\Delta_{lk}}
\end{align}$$

China shock (Autor, Dorn and Hanson, 2013):
- $z_{lk}$: location, $l$, and industry, $k$, composition
- $\Delta_{lk}$: location, $l$, and industry, $k$, growth in imports from China
- $\Delta_{k}$: industry $k$ growth of imports from China

---
# Other Examples

$$\begin{align}
B_{l} &= \sum_{k=1}^{K} z_{lk} \Delta_{k},\\
\Delta_{lk} &= \Delta_{k} + \tilde{\Delta_{lk}}
\end{align}$$

Immigrant enclave (Altonji and Card, 1991):
- $z_{lk}$: share of people from foreign country $k$ living in current location $l$ (in base period)
- $\Delta_{lk}$: growth in the number of people from $k$ to $l$
- $\Delta_{k}$: growth in people from $k$ nationally

---
# Other Examples

$$\begin{align}
B_{l} &= \sum_{k=1}^{K} z_{lk} \Delta_{k},\\
\Delta_{lk} &= \Delta_{k} + \tilde{\Delta_{lk}}
\end{align}$$

Market size and demography (Acemoglu and Linn, 2004):
- $z_{lk}$: spending share on drug $l$ from age group $k$
- $\Delta_{lk}$: growth in spending of group $k$ on drug $l$
- $\Delta_{k}$: national growth in spending of group $k$ (e.g., due to population aging)


---
# Bartik Instruments

- Goldsmith-Pinkham, Sorkin, and Swift (2020) show that using $B_{l}$ as an instrument is equivalent to using local industry shares, $z_{lk}$, as IVs
- Can decompose Bartik-style IV estimates into weighted combination of estimates where each share is an instrument (Rotemberg weights)
- Borusyak, Hull, and Jaravel (2022), ReStud, instead focus on situation in which the shifts are exogenous and the shares are potentially endogenous
- Borusyak and Hull (2021), Econometrica (maybe), provide general approach when using exogenous shifts (recentering as in homework)


**key:** literature was vague as to the underlying source of variation in the instrument...recent papers help in understanding this (and thus defending your strategy)

---
# Shift-Share (focusing on the shift)

- Did ACA medicaid expansion affect insurance status?
- Construct "simulated" IV (dummy for whether person $i$ is newly eligible given state expansion)
- Instrument can be thought of as, $z_{is} = f(x_{i}, e_{s})$
- Want to separate variation due to $e_{s}$ (the state policy changes) from variation in demographics, $x_{i}$
    - Identify $p(x)$, probability of eligibility on average across other states' laws
    - Recenter, $\tilde{z}_{is} = z_{is} - p(x)$. Can also just control for $p(x)$ in regression.

