---
title: "Class Schedule"
citeproc: true
bibliography: syllabus/BibTeX_Library.bib
csl: syllabus/chicago-syllabus-no-bib.csl
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(googledrive)
library(dplyr)
library(tibble)
library(pander)
```

```{r dates_bib, include=FALSE}
start <- as.Date("2022-08-24")
advdate <- function(adv) {
 twed <- start + 7*floor(adv/2)
 tmon <- twed-2
 tmon <- format(tmon, format="%m/%d")
 twed <- format(twed, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tday <- ifelse( (adv %% 2) == 0, tmon,twed)
 tmp <- paste("Class ",zadv,sep='',", ", tday)
 return(tmp)
}
```


Below is a preliminary course outline for this semester. For a more complete reading list each day as well as more detail on course policies, please see the full syllabus [here](syllabus/syllabus-771.pdf).


## Module 0: Kicking things off!

### `r advdate(1)`: Introduction to the economics of healthcare**

**General health econ references:**<br>

  - @arrow1963
  - @dranove2000
  - @evans2000
  - @handel2021
  - @gaynor2015jel


### `r advdate(2)`: Introduction to causal inference**

  - Chapters 1-5 of [Causal Inference: The Mixtape](https://mixtape.scunning.com/)
  - Chapters 1-5, 10, 13, and 14 of [The Effect: An Introduction to Research Design and Causality](https://theeffectbook.net/)
  

## Module 1: Hospitals

### `r advdate(3)`: Background and institutional details**

**Supplemental reading:**<br>

  - @newhouse1970
  - @bayindir2012
  - @sloan2000
  - @dranove1988
  - @pauly1978
  - @duggan2000
  - @horwitz2009
  - @frank1991
  - @gruber1994
  - @sloan2001
  - @deneffe2002
  - @duggan2002
  - @kessler2002
  - @gaynor2003
  - @david2009
  - @chang2017

### `r advdate(4)` - `r advdate(5)`: Hospitals and financial incentives

**Primary reading:**<br>

  - @dafny2005
  - @eliason2018
  - @batty2017
  - @lee2020

**Supplemental reading:**<br>

  - @dranove1987rand
  - @cutler1995
  - @ellis1996
  - @silverman2004
  - @carroll2018
  - @cook2020
  - @desai2018
  - @gowrisankaran2019
  - @gupta2021


### `r advdate(6)` - `r advdate(8)`: Application

  - **Question:** How do hospitals respond to pay for performance programs?
  - **Research design:** Difference-in-differences

**Primary reading:**<br>

  - @gupta2021
  - @callaway2021
  - @rambachan2022
  - @roth2022
  - @arkhangelsky2021
  
**Supplemental reading:**<br>


## Module 2: Physicians

### `r advdate(8)`: Background and institutional details**

**Supplemental reading:**<br>

  - @mcguire2000
  - @grossman1983
  - @wennberg1973
  - @wennberg2003
  - @skinner1997
  - @finkelstein2016
  - @molitor2018
  

### `r advdate(9)`: Physician agency, financial incentives, and referrals

**Primary reading:**<br>

  - @clemens2014
  - @zeltzer2020

**Supplemental reading::**<br>

  - @baker2016
  - @gruber1996
  - @afendulis2007
  - @iizuka2012
  - @whaley2021
  - @young2021
  - @walden2016
  - @richards2022
  - @baker2016
  - @chernew2021
  
  
## `r advdate(10)`: Physician learning

**Primary reading:**<br>
  
  - @crawford2005
  - @sarsons2018

**Supplemental reading:**<br>

  - @coscelli2004
  - @ferreyra2011
  - @chan2013
  - @dickstein2018
  - @gong2018
  - @johnson2011
  - @ho2002


### `r advdate(11)` - `r advdate(13)`: Application

  - **Question:** Physician effort and hospital ownership
  - **Research design:** Instrumental variables

**Primary reading:**<br>
  
  - @dranove2019
  - @borusyak2021restud

**Supplemental reading:**<br>


### `r advdate(14)`: Fall break

## Module 3: Information disclosure

### `r advdate(15)`: Patient decision-making

**Primary reading:**<br>

  - @ketcham2012
  - @gaynor2016

**Supplemental:**<br>

  - @farrell2007
  - @abaluck2011


### `r advdate(16)`: Inertia in health insurance choice

**Primary reading:**<br>

  - @handel2013
  - @ericson2014


### `r advdate(17)`: Effects of quality disclosure

**Primary reading:**<br>

  - @dranove2003
  - @dafny2008
  - @kolstad2013

**Supplemental:**<br>

  - @chernew1998
  - @scanlon2002
  - @wedig2002
  - @jin2006
  - @dranove2010
  - @reid2013
  - @darden2015
  - @mccarthy2017
  - @epstein2010  



### `r advdate(18)`: Effects of price disclosure

**Primary reading:**<br>

  - @chernew2021
  - @grennan2020

### `r advdate(19)` and `r advdate(20)`: Application

  - **Question:** Consumer inertia and insurer premiums
  - **Research design:** Regression discontinuity

**Primary reading:**<br>
  
  - @ericson2014

**Supplemental reading:**<br>


## Module 4: Healthcare competition and industrial organization

### `r advdate(21)`: Competition in price and quality**

**Primary reading:**<br>

  - @dranove1992
  - @fan2013
  - @lewis2016

**Supplemental:**<br>

  - @spence1975
  - @mussa1978
  - @hausman1994
  - @dana2011

### `r advdate(22)`: Measuring hospital markets and market power

**Primary reading:**<br>

  - @dranove2016
  - @everson2019

**Supplemental:**<br>

  - @dranove1990
  - @capps2003



### `r advdate(23)`: "Structure-Conduct-Performance"

**Primary reading:**<br>

  - @kessler2000
  - @cooper2019

**Supplemental:**<br>

  - @dranove1992
  - @lynk1995
  - @keeler1999


### `r advdate(24)`: Bargaining framework

**Primary reading:**<br>

  - @ho2017
  - @clemens2017

**Supplemental:**<br>

  - @bresnahan1991
  - @reiss2007
  - @abraham2007
  - @gowrisankaran2015
  - @lewis2015


### `r advdate(25)`: Effects of mergers (or closures)

**Primary reading:**<br>

  - @craig2021
  - @eliason2020  

**Supplemental:**<br>

  - @vita2001
  - @gaynor2003
  - @lindrooth2003
  - @gaynor2012rwjf
  - @dafny2009
  - @gaynor2013


### `r advdate(26)`: No meeting today. Class rescheduled for December 2


### `r advdate(27)`: Out-of-market mergers (**virtual class**)

**Primary reading:**<br>

  - @dafny2019
  - @schmitt2018

### `r advdate(28)`: Vertical integration

**Primary reading:**<br>

  - @luco2020
  - @koch2021

**Supplemental:**<br>

  - @dranove2019
  - @koch2017
  - @lin2021
  - @burns2002
  - @ciliberto2006
  - @cuellar2006
  - @baker2016
  - @neprash2015
  - @capps2018
  - @konetzka2018


### 12/02 (rescheduled from 11/21) and `r advdate(30)`: Application

  - **Question:** Sensitivity of demand estimates to market definition
  - **Research design:** Structural demand estimation

**Primary reading:**<br>
  
  - @everson2019

**Supplemental reading:**<br>



## Module 5: Health insurance markets (we won't get to this)

### Insurance markets and competitiveness

**Primary reading:**<br>

  - @dafny2010
  - @dafny2012
  - @bundorf2012

**Supplemental:**<br>

  - @town2003
  - @einav2010
  - @einav2010qje
  - @starc2014


### Adverse selection

**Primary reading:**<br>

  - @cutler1998
  - @decarolis2017

**Supplemental:**<br>

  - @ackerloff1970
  - @rothschild1976
  - @frank2000
  - @einav2011


### Moral hazard

**Primary reading:**<br>

  - @einav2013

**Supplemental:**<br>

  - @manning1996
  - @finkelstein2014
  - @einav2018


### Managed competition

**Primary reading:**<br>

  - @curto2021
  - @einav2015

**Supplemental:**<br>

  - @song2013
  - @cabral2018
  - @stockley2014
  - @duggan2016
  - @pelech2018






