---
title: "More on Instrumental Variables"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Econ 771, Fall 2022
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{tikz}
  - \usepackage{pgf}
  - \usepackage{pgfarrows}
  - \usepackage{pgfnodes}
  - \usepackage{pgfautomata}
  - \usepackage{pgfheaps}
  - \usepackage[absolute,overlay]{textpos}  
  - \usepackage{mathtools}
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
bibliography: '../syllabus/BibTeX_Library.bib'      
---
class: inverse, center, middle



<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, fixest, readstata13,
               xaringanExtra, webshotm, ggthemes, gganimate, plotly, ivpack, ivreg)
set.seed(1234)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


# Problem of Weak Instruments

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Why we care about instrument strength

Recall our schooling and wages equation, $$y = \beta S + \epsilon$$. Bias in IV can be represented as:

$$Bias_{IV} \approx \frac{Cov(S, \epsilon)}{V(S)} \frac{1}{F+1} = Bias_{OLS} \frac{1}{F+1}$$

- Bias in IV may be close to OLS, depending on instrument strength
- **Bigger problem:** Bias could be bigger than OLS if exclusion restriction not *fully* satisfied
- Over-reliance on "rules of thumb", as seen in [Anders and Kasy (2019)](https://www.aeaweb.org/articles?id=10.1257/aer.20180310)



---
# Testing strength of instruments

.pull-left[
**Single endogenous variable**
- F-test of instruments (rule of thumb critical value of 10)
- Partial $R^2$
]

.pull-right[
**Many endogenous variables**
- More complicated
]


---
class: inverse, center, middle

# Interpreting Instrumental Variables Results

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# LATE and IV Interpretation

- With monotonicity assumption (all those affected by the instrument are affected in the same "direction")
- In the face of **heterogeneous treatment effects**, IV provides a "Local Average Treatment Effect"
- LATE: Effect of treatment among those affected by the instrument (compliers only)
- Why does this matter? Let's discuss Medicaid expansion in Oregon



---
class: inverse, center, middle

# "Modern" IV

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Machine learning and IV