---
title: "Basics of Demand Estimation"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Econ 771, Fall 2022
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{tikz}
  - \usepackage{pgf}
  - \usepackage{pgfarrows}
  - \usepackage{pgfnodes}
  - \usepackage{pgfautomata}
  - \usepackage{pgfheaps}
  - \usepackage[absolute,overlay]{textpos}  
  - \usepackage{mathtools}
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
bibliography: '../syllabus/BibTeX_Library.bib'      
---
class: inverse, center, middle


<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, fixest, readstata13,
               xaringanExtra, rdrobust, ggthemes, gganimate, plotly, here, hrbrthemes,
               scales, gapminder, gifski, png, tufte, OECD, ggrepel, dtplyr, data.table, kableExtra, 
               cobalt, haven, magick)
set.seed(1234)
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble(rgb(0.9, 0.5, 0.5))
```


# Discrete Choice

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Setup

Indirect utility of person $i$, $$u_{ij} = x_{ij}\beta + \epsilon_{ij},$$ where $x_{ij}$ denotes person (and perhaps product) characteristics and $\epsilon_{ij}$ denotes an error term. 

- Standard logit: one choice, $j=0,1$
- Multinomial logit: many possible choices, $j=0,1,...,J$


---
# Logit terminology

A few different terms for very similar models:

- Multinomial Logit
- Conditional Logit
- "Mixed" Logit

---
# Multinomial 

- Individual covariates only
- Alternative-specific coefficients

--
$$u_{ij}=x_{i}\beta_{j} + \epsilon_{ij},$$ such that $$p_{ij} = \frac{e^{x_{i}\beta_{j}}}{\sum_{k} e^{x_{i}\beta_{k}}}$$


---
# Conditional

Allow for alternative-specific regressors, such that $$u_{ij}=x_{ij}\beta + \epsilon_{ij}$$


---
# Mixed

Allow for individual and alternative-specific regressors, such that $$u_{ij}=x_{ij}\beta + w_{i} \gamma_{j} + \epsilon_{ij}$$

<br>
--
*but* people sometimes use "mixed" to refer to random-coefficients logit


---
# Does it matter?

These are really all the same and it's just a matter of specification (e.g., interact individual covariates with product characteristics or with product dummies). I'll refer to them as "multinomial" logit.


---
# The Indepenence of Irrelevant Alternatives

Fundamental issue with logit models...the ratio of choice probabilities for $j$ and $k$ does not depend on any other alternatives: $$\frac{P_{ij}}{P_{ik}} = \frac{e^{V_{ij}}}{e^{V_{ik}}}.$$

---
# Relaxing IIA

- This is really an omitted variables problem...with enough interactions, we can allow for a sufficiently rich substitution pattern
- Alternatively, relax assumptions on the error term with nested logit or random-coefficient logit (or multinomial probit)

<!-- New Section -->
---
class: inverse, center, middle

# Discrete Choice with Market Data

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Setup

Utility of individual $i$ from selecting product $j$ is $$U_{ij}=\delta_{j}+\epsilon_{ij},$$
where $\delta_{j}=x_{j}\beta + \xi_{j}$, and $\xi_{j}$ represents the mean level of utility derived from unobserved characteristics.

---
# Estimating with market shares

Goal is to find $\hat{\delta}$ to statisfy moment condition, $$\frac{1}{J}\sum_{j} (\hat{\delta}_{j}-x_{j}\beta)z_{j}=0.$$

In standard logit, $s_{j}=e^{\delta_{j}}/\sum e^{\delta_{j}}$, and $\delta_{j}$ then follows directly from taking logs and subtracting the outside share (with the normalization of $\delta_{0}=0$, which yields the estimating equation $$\ln(s_{j}) - \ln(s_{0}) = x_{j}\beta + \xi_{j}$$

---

# Estimating with market shares

- Standard logit imposes cross-price elasticities that are proportional to market shares (limited substitution patterns)
- Relax with nested logit or random-coefficients logit
- For details, see Berry (1994) or come talk to me

