\documentclass[11pt,]{article}
\usepackage[margin=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}
\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center
\newcommand{\blankline}{\quad\pagebreak[2]}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}} 
\usepackage{longtable,booktabs}

\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}

\titleformat*{\subsubsection}{\normalsize\itshape}

\usepackage{titling}
\setlength{\droptitle}{-.25cm}

%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines 

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{lastpage}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0.0pt} 
\lhead{}
\chead{}
\rhead{\footnotesize Econ 771: Health Economics II -- 8/24/2022 -
12/6/2022}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\fancypagestyle{firststyle}
{
\renewcommand{\headrulewidth}{0pt}%
   \fancyhf{}
   \fancyfoot[C]{\small \thepage/\pageref*{LastPage}}
}

%\def\labelitemi{--}
%\usepackage{enumitem}
%\setitemize[0]{leftmargin=25pt}
%\setenumerate[0]{leftmargin=25pt}




\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
}
\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={Econ 771: Health Economics II},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls


\setcounter{secnumdepth}{0}





\usepackage{setspace}

\title{Econ 771: Health Economics II}
\author{Ian McCarthy}
\date{8/24/2022 - 12/6/2022}


\begin{document}  

		\maketitle
		
	
		\thispagestyle{firststyle}

%	\thispagestyle{empty}


	\noindent \begin{tabular*}{\textwidth}{ @{\extracolsep{\fill}} lr @{\extracolsep{\fill}}}


E-mail: \texttt{\href{mailto:ian.mccarthy@emory.edu}{\nolinkurl{ian.mccarthy@emory.edu}}} & Web: \href{http://econ771f22.classes.ianmccarthyecon.com}{\tt econ771f22.classes.ianmccarthyecon.com}\\
Office Hours: MoWe 01:00-2:30 p.m.  &  Class Hours: MoWe 10:00-11:15
a.m\\
Office: PAIS 573  & Class Room: North Decatur Building, Room 109\\
	&  \\
	\hline
	\end{tabular*}
	
\vspace{2mm}
	


\hypertarget{course-description}{%
\section{Course Description}\label{course-description}}

This course explores the industrial organization of healthcare markets
in the U.S. We will focus on the following areas: hospital ownership and
production, physician agency, information disclosure, and hospital
competition. If we have time, we will briefly cover some issues on
health insurance markets. The class is effectively designed as a
combination of empirical IO and causal inference, with applications to
healthcare. As such, we will also examine several econometric tools and
causal inference identification strategies. These methods will be
introduced as needed throughout the course.

Supply-side health economics is a rapidly growing field with many new
developments, particularly in the areas of bargaining in two-sided
markets and physician learning. Some of these recent developments use
tools from network analysis and machine learning, which we unfortunately
do not have time to cover in this course. I've also chosen specific
topics that overlap most with my own research --- the upside here is
that I can speak somewhat confidently about the literature and empirical
studies in this area, but the downside is that some very interesting
areas of health economics are not studied. For example, we will largely
ignore issues of the prescription drug market, medical devices, and
physician labor supply. My hope is that the content that we do cover
will provide a springboard for those interested in these other important
areas.

\hypertarget{learning-outcomes}{%
\section{Learning Outcomes}\label{learning-outcomes}}

I have five central goals for this course:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Synthesize the current literature in each of the main areas of health
  economics covered in this class
\item
  Apply standard causal inference techniques in the area of healthcare
\item
  Provide constructive criticism of academic work in this area
\item
  Organize real-world data with current workflow and versioning tools
\item
  Develop your own preliminary research in some area of health economics
\end{enumerate}

There are more specific learning outcomes for each module described on
the relevant module's page of our class website, accessible
\href{https://imccart.github.io/Econ-771/}{here}.

\hypertarget{text-software-and-class-materials}{%
\section{Text, Software, and Class
Materials}\label{text-software-and-class-materials}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Readings:} As an elective PhD course, we will rely on academic
  papers from the reading list below. I expect everyone to read the
  papers in advance and come to class with questions on the study's
  contribution, empirical techniques, identification strategies, and
  datasets used. My goal with each paper is to discuss the analysis in
  detail. As such, the primary reading list is perhaps shorter than a
  standard PhD course. I've provided supplemental reading in each
  section for those interested in additional readings in a specific
  area.
\item
  \textbf{Software:} For almost everything in this class, I'll use
  \texttt{R}, but you are free to use whatever software you're most
  comfortable with. I encourage you to use \texttt{R}, \texttt{Stata},
  or \texttt{Python} simply because these are the most common programs
  used in practice right now. You'll also need to have a basic working
  knowledge of Git and GitHub. If you're new to these tools, take a look
  at \href{https://grantmcdermott.com/}{Grant McDermott's} notes on
  \href{https://github.com/uo-ec607/lectures}{Data Science for
  Economists} as well as \href{https://jennybryan.org/}{Jenny Bryan's}
  online reference book, \href{https://happygitwithr.com/}{Happy Git and
  GitHub with R}.
\item
  \textbf{Accessing Data:} For any in-class activities, I will house all
  of the code and links to the data on our class website. For any
  real-life datasets, I will also point you to other GitHub repositories
  where available, and I'll make some data available via OneDrive.
  Details will be posted on Canvas and in class.
\item
  \textbf{Logistics:} For day-to-day communication, grades, and other
  private information (such as Zoom meeting links if needed), I'll use
  \emph{Canvas}. I'll post all other materials to our class website.
\item
  \textbf{Slides and Notes:} Any presentations will be made available on
  our class website prior to any given class.
\end{enumerate}

\hypertarget{course-policies}{%
\section{Course Policies}\label{course-policies}}

Various policies for this course are described below. Basically, let's
all work to be good citizens and take seriously our various roles as a
student, teacher, friend, colleague, human, etc.

\hypertarget{academic-integrity}{%
\subsection{Academic integrity}\label{academic-integrity}}

The Emory University Honor Code is taken seriously and governs all work
in this course. Details about the Honor Code are available in the Laney
Graduate School Handbook and available online
\href{https://www.gs.emory.edu/academics/policies-progress/conduct.html}{here}.
By taking this course, you affirm that it is a violation of the code to
plagiarize, to deviate from the instructions about collaboration on work
that is submitted for grades, to give false information to a faculty
member, and to undertake any other form of academic misconduct. You also
affirm that if you witness others violating the code you have a duty to
report them to the honor council.

\hypertarget{accessibility-services}{%
\subsection{Accessibility services}\label{accessibility-services}}

If you anticipate issues related to the format or requirements of this
course, please meet with me. I would like us to discuss ways to ensure
your full participation in the course. If you determine that
accommodations are necessary, you may register with Accessibility
Services at (404)727-9877 or via e-mail at
\href{mailto:accessibility@emory.edu}{\nolinkurl{accessibility@emory.edu}}.
To register with OAS, students must self identify and initiate contact
with the OAS office.

\hypertarget{communication}{%
\subsection{Communication}\label{communication}}

I will post regular announcements to the class on \emph{Canvas}, so
please set up your notifications on \emph{Canvas} accordingly. I will
also use \emph{Canvas} to post all grades and any other information that
needs to stay in the class (like our \emph{Zoom} meeting link for
virtual meetings, if needed). All other course materials will be
available on our class website,
\href{https://econ771f22.classes.ianmccarthyecon.com/}{econ771f22.classes.ianmccarthyecon.com/}.

Please feel free to reach out to me for any reason. I generally respond
to all e-mails within 24 hours.

\hypertarget{office-hours}{%
\subsection{Office Hours}\label{office-hours}}

My office is in room 573 of the PAIS building. This is where I'll be for
official office hours; however, I'm happy to meet outside of normal
office hours as well. Please feel free to schedule another time to meet
by following the link to select a time that works for you,
\url{https://mccarthy-meetings.youcanbook.me}.

\hypertarget{attendance}{%
\subsection{Attendance}\label{attendance}}

While there is no official ``attendance'' credit, everyone is expect to
attend all class sessions. Given our small class, it is very important
that we are all present and engaged.

\hypertarget{assignments-and-grading}{%
\section{Assignments and Grading}\label{assignments-and-grading}}

There are four main assignments throughout the semester, along with a
participation component. I describe each assignment below, with more
detail provided on our \href{https://imccart.github.io/Econ-771/}{class
website}. Note that I \textbf{do not} expect or want anyone to treat
these assignments as entirely separate. By that, I mean you should try
to identify datasets from the exercises that you can use in your
replication project, and you should try to select papers to present that
use those data. At a minimum, you should present the paper that you will
be replicating. With a little bit of up-front planning, you can create a
lot of overlap across assignments, which will make your life much
easier.

\hypertarget{participation}{%
\subsection{Participation}\label{participation}}

Everyone will benefit much more from the class if we are all actively
engaged. As a small incentive to stay engaged, I'll keep track of those
who participate in our discussion each day. Any amount of participation
(questions or comments) will be worth one point for that class. I'll
tally everyone's participation points at the end of the semester.

\hypertarget{empirical-exercises}{%
\subsection{Empirical exercises}\label{empirical-exercises}}

Each course module will have an applied component where we spend some
time replicating analyses from selected papers based on real-world data.
These will require some of your time outside of class to get the data in
working order and implement the relevant identification strategy and
econometric estimator. This empirical work should heavily complement
your draft paper. The due dates for all empirical exercises are
indicated in the schedule below. In general, we have one empirical
assignment due at the end of each module.

For each assignment, your final product should consist of a discussion
of the data sources, steps taken to clean the data and replicate the
final analytic sample, a summary of your model and estimator, and your
results. You must turn in your assignment on \emph{GitHub}, including
your code files (Stata, R, Python, SAS, etc.), all figures/tables, and
of course the final write-up. I expect the final product to be similar
to an academic paper in terms of the quality of figures, tables, and
general formatting.

Please organize your folders in a useful way. The way I organize things
(though certainly not the only way to do it) is to keep a folder for
each new project and named accordingly. I typically have the following
subfolders:

\begin{itemize}
\item
  \emph{Data}: This is where I keep the raw data files and any
  additional data files I create as part of my analysis. I also keep a
  ``Research Data'' folder on my computer that has raw data that I
  access regularly, in which case the ``Data'' folder in any given
  project includes symbolic links to the original raw data.
\item
  \emph{Data Code}: This is where I keep the code for data management
  (merging, clean-up, wrangling, etc.). Final analytic datasets created
  by this code go to the ``Data'' subdirectory.
\item
  \emph{Analysis}: This is where I keep my analysis code files and log
  files (if relevant).
\item
  \emph{Results}: Output from the analysis code files. I tend to
  separate this into two subfolders - one for tables and one for
  figures.
\item
  \emph{Papers}: If I'm writing the paper in \texttt{R\ Markdown}, this
  is where I keep all of the markdown files. With most co-authored work,
  though, we write in Overleaf. In that case, I'll still download
  different versions of the paper into this folder when necessary (e.g.,
  for different journal submissions)
\item
  \emph{Presentation}: This is where I keep my slides and underlying
  code files.
\end{itemize}

It's good to start developing some organization practices that work best
for you. It's extremely easy to forget what you were doing on a project
once you have several things going at once, especially when you wait for
6-8 months after submitting a paper for publication. The last thing you
want is to not be able to replicate your own work!

\hypertarget{presentations}{%
\subsection{Presentations}\label{presentations}}

You will present three papers throughout the course of the semester. Any
paper listed as ``Primary Reading'' in the class schedule can be
selected for presentation, with the exception of papers indicated by **
. Please inform me of your selected papers no later than \textbf{Friday,
September 2}. Please also note your papers and class dates on the Google
Sheet
\href{https://docs.google.com/spreadsheets/d/1d6IVg-khbreLKobiczNwEOT9gjPNNykrFEtc2rXdJ4I/edit?usp=sharing}{here}.

Each presentation should be no more than 30 minutes, including questions
and brief discussion. The presentation should follow a standard
conference setup, with a brief introduction/motivation, a very brief
discussion of the literature and some context of the paper, discussion
of the data, empirical analysis, and results.

Note that a presentation is not just a re-hashing of the paper in slide
form. A good academic presentation should have as little information as
possible on each slide, and the content on the slides doesn't
necessarily need to follow that of the paper. For example, in a
real-time environment, it is much easier to move between different
aspects of the empirical analysis and data.

In addition to the presentation, please send me your slides in advance.
For people in the economics department, I expect the slides to be
completed in Beamer (LaTeX), \texttt{R\ Markdown}, or Python. Others can
use PowerPoint or some other presentation tool as they see fit. My only
recommendation with those other programs is that, for some of them like
Prezi or Powtoon, it is easy for the presentation to become distracting.
Slides should complement your presentation and not replace your
presence!

Grades for each presentation will be based 50\% on the slides and 50\%
on the actual presentation (delivery, clarity/organization, and
content). Additional details and a grading rubric are available on the
assignments page of our class website.

\hypertarget{literature-review-2nd-year-students}{%
\subsection{Literature review (2nd year
students)}\label{literature-review-2nd-year-students}}

The literature review can be on any health economics topic of your
choice, subject to my approval. There is no specific page requirement.
10 double-spaced pages might be a good target, but an
efficiently-written paper could be shorter, while a student wishing to
use the paper as a springboard to a dissertation may choose to write
more. The paper will mostly consist of discussions of prior research,
but should end with discussions of three open questions in the
literature plus a proposed strategy for answering at least one of these
questions. The paper is due by \textbf{Friday, November 18}.

\hypertarget{draft-paper-3rd-year-students}{%
\subsection{Draft paper (3rd year
students)}\label{draft-paper-3rd-year-students}}

The paper requirement is more extensive for 3rd year students. In this
case, your papers should include not only an extensive literature review
but also a preliminary empirical analysis, including a discussion of the
data you're using, the construction of your sample, your identification
strategy, econometric model and estimator, and some preliminary results.
There is again no specific page requirement, but I expect at least 20
double-spaced pages to appropriately discuss your topic, context, data,
and early results. Your draft paper should end with an outline of
additional analyses you hope to run (i.e., robustness/sensitivity
analyses, a discussion of different mechanisms of interest,
policy-relevant heterogeneities in your estimated effects, etc.). The
paper is due by \textbf{Friday, November 18}.

\hypertarget{due-dates}{%
\subsection{Due dates}\label{due-dates}}

This section is just to highlight the most important dates on which an
assignment is due. These dates are repeated within the description of
each assignment above. Note that you can always turn assignments in
prior to the official due date. Late assignments will receive an
automatic 10\% reduction in the grade for each day the assignment is
turned in after the due date.

\begin{itemize}
\tightlist
\item
  \textbf{September 2}: Note the papers you plan to present on the
  Google Sheet
  \href{https://docs.google.com/spreadsheets/d/1d6IVg-khbreLKobiczNwEOT9gjPNNykrFEtc2rXdJ4I/edit?usp=sharing}{here}.
  Only one student can present each paper, so this is first-come
  first-served.
\item
  \textbf{Friday, November 18}: Due date for literature review or draft
  paper.
\item
  Empirical exercises due one week after the end of each module
\end{itemize}

\hypertarget{final-grades}{%
\subsection{Final grades}\label{final-grades}}

\begin{itemize}
\item
  \textbf{10\%} for participation (out of 25 possible points)
\item
  \textbf{40\%} for empirical exercises (10\% each)
\item
  \textbf{30\%} for presentations of selected papers (10\% each)
\item
  \textbf{20\%} for literature review (2nd year students) or draft paper
  (3rd year students)
\end{itemize}

Letter grades will be assigned at the end of the course based on total
score achieved: (A = 100-93\%, A- = 92.99-90\%, B+ = 89.99-87\%, B =
86.99-83\%, B- = 82.99-80\%, C+ = 79.99-77\%, C = 76.99-73\%, C- =
72.99-70\%, D+ = 69.99-67\%, D = 66.99-60\%, F = 60\% or less)

\newpage

\hypertarget{class-schedule}{%
\section{Class Schedule}\label{class-schedule}}

Below is a preliminary outline of specific topics and readings
throughout the semester. Based on our collective interests, discussions,
and timing, individual papers and order may change somewhat, but any
such changes will be announced in advance. For each class, one of us
will present the ``primary reading'' papers, followed by a general
discussion of the papers and topic among the class. Papers listed as
``supplemental'' represent other relevant papers in that area that we
won't have time to cover directly in class. Classes listed with **
denote ``lecture'' days. I'll handle presentations and manage any
in-class activities on those days.

Note that we have two scheduled virtual classes, one for Monday, 9/19,
and another for Wednesday, 11/23. We also need to reschedule our class
for Monday, 11/21. I've tentatively rescheduled that for Friday,
December 2.

\hypertarget{module-0-kicking-things-off}{%
\subsection{Module 0: Kicking things
off!}\label{module-0-kicking-things-off}}

\hypertarget{class-01-0824-introduction-to-the-economics-of-healthcare-and-causal-inference}{%
\subsubsection{Class 01, 08/24: Introduction to the economics of
healthcare and causal
inference**}\label{class-01-0824-introduction-to-the-economics-of-healthcare-and-causal-inference}}

\textbf{General health econ references:}

\begin{itemize}
\tightlist
\item
  Kenneth J. Arrow {``Uncertainty and the {Welfare} {Economics} of
  {Medical} {Care},''} \emph{The American Economics Review} 53, no. 5
  (1963).
\item
  David Dranove and Mark A Satterthwaite {``The Industrial Organization
  of Health Care Markets,''} \emph{Handbook of Health Economics} 1
  (2000): 1093--1139.
\item
  William N Evans, Helen Levy, and Kosali I Simon {``Data Watch:
  {Research} Data in Health Economics,''} \emph{Journal of Economic
  Perspectives} 14, no. 4 (2000): 203--216.
\item
  Ben Handel and Kate Ho {``The Industrial Organization of Health Care
  Markets,''} in \emph{Handbook of {Industrial} {Organization}}, ed.
  Kate Ho, Ali Hortaçsu, and Alessandro Lizzeri, vol. 5, Handbook of
  {Industrial} {Organization}, {Volume} 5 (Elsevier, 2021), 521--614.
\item
  M Gaynor, K Ho, and R Town {``The {Industrial} {Organization} of
  {Health} {Care} {Markets},''} \emph{Journal of Economic Literature}
  47, no. 2 (2015): 235--284.
\end{itemize}

\textbf{General causal inference references:}

\begin{itemize}
\tightlist
\item
  Chapters 1-5 of \href{https://mixtape.scunning.com/}{Causal Inference:
  The Mixtape}
\item
  Chapters 1-5, 10, 13, and 14 of \href{https://theeffectbook.net/}{The
  Effect: An Introduction to Research Design and Causality}
\end{itemize}

\hypertarget{module-1-hospitals}{%
\subsection{Module 1: Hospitals}\label{module-1-hospitals}}

\hypertarget{class-02-0829-background-and-institutional-details}{%
\subsubsection{Class 02, 08/29: Background and institutional
details**}\label{class-02-0829-background-and-institutional-details}}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Joseph P Newhouse {``Toward a Theory of Nonprofit Institutions: {An}
  Economic Model of a Hospital,''} \emph{American Economic Review}
  (1970): 64--74.
\item
  Esra Eren Bayindir {``Hospital Ownership Type and Treatment
  Choices,''} \emph{Journal of Health Economics} 31, no. 2 (2012):
  359--370.
\item
  Frank A Sloan {``Not-for-Profit Ownership and Hospital Behavior,''}
  \emph{Handbook of Health Economics} 1 (2000): 1141--1174.
\item
  David Dranove {``Pricing by Non-Profit Institutions: The Case of
  Hospital Cost-Shifting,''} \emph{Journal of Health Economics} 7, no. 1
  (1988): 47--57.
\item
  Mark V Pauly {``Medical Staff Characteristics and Hospital Costs,''}
  \emph{Journal of Human Resources} (1978): 77--111.
\item
  Mark Duggan {``Hospital {Ownership} and {Public} {Medical}
  {Spending},''} \emph{Quarterly Journal of Economics} 115, no. 4
  (2000): 1343--1373.
\item
  Jill R Horwitz and Austin Nichols {``Hospital Ownership and Medical
  Services: Market Mix, Spillover Effects, and Nonprofit Objectives,''}
  \emph{Journal of Health Economics} 28, no. 5 (2009): 924--937.
\item
  Richard G Frank and David S Salkever {``The Supply of Charity Services
  by Nonprofit Hospitals: Motives and Market Structure,''} \emph{RAND
  Journal of Economics} (1991): 430--445.
\item
  Jonathan Gruber {``The Effect of Competitive Pressure on Charity:
  {Hospital} Responses to Price Shopping in {California},''}
  \emph{Journal of Health Economics} 13, no. 2 (1994): 183--211.
\item
  Frank A Sloan et al. {``Hospital Ownership and Cost and Quality of
  Care: Is There a Dime's Worth of Difference?''} \emph{Journal of
  Health Economics} 20, no. 1 (2001): 1--21.
\item
  Daniel Deneffe and Robert T Masson {``What Do Not-for-Profit Hospitals
  Maximize?''} \emph{International Journal of Industrial Organization}
  20, no. 4 (2002): 461--492.
\item
  Mark Duggan {``Hospital Market Structure and the Behavior of
  Not-for-Profit Hospitals,''} \emph{RAND Journal of Economics} (2002):
  433--446.
\item
  Daniel P Kessler, Mark B McClellan, et al. {``The {Effects} of
  {Hospital} {Ownership} on {Medical} {Productivity},''} \emph{RAND
  Journal of Economics} 33, no. 3 (2002): 488--506.
\item
  Martin Gaynor and William B Vogt {``Competition Among {Hospitals},''}
  \emph{RAND Journal of Economics} (2003): 764--785.
\item
  Guy David {``The Convergence Between for-Profit and Nonprofit
  Hospitals in the {United} {States},''} \emph{International Journal of
  Health Care Finance and Economics} 9, no. 4 (2009): 403--428.
\item
  Tom Chang and Mireille Jacobson \emph{What Do {Nonprofit} {Hospitals}
  {Maximize}? {Evidence} from {California}'s {Seismic} {Retrofit}
  {Mandate}}, Working \{Paper\} (Case Western Reserve University, 2017).
\end{itemize}

\hypertarget{class-03-0831-hospitals-and-financial-incentives-part-i}{%
\subsubsection{Class 03, 08/31: Hospitals and financial incentives, part
I}\label{class-03-0831-hospitals-and-financial-incentives-part-i}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Leemore S. Dafny {``How {Do} {Hospitals} {Respond} to {Price}
  {Changes}?''} \emph{American Economic Review} 95, no. 5 (2005):
  1525--1547.
\item
  Paul J Eliason et al. {``Strategic Patient Discharge: {The} Case of
  Long-Term Care Hospitals,''} \emph{American Economic Review} 108, no.
  11 (2018): 3232--65.
\item
  Michael Batty and Benedic Ippolito {``Financial Incentives, Hospital
  Care, and Health Outcomes: {Evidence} from Fair Pricing Laws,''}
  \emph{American Economic Journal: Economic Policy} 9, no. 2 (2017):
  28--56.
\item
  Ajin Lee {``How Do Hospitals Respond to Managed Care? {Evidence} from
  at-Risk Newborns,''} \emph{Journal of Public Economics} 184 (2020):
  104130.
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  David Dranove {``Rate-Setting by Diagnosis Related Groups and Hospital
  Specialization,''} \emph{The RAND Journal of Economics} (1987):
  417--427.
\item
  David M Cutler {``The {Incidence} of {Advserse} {Medical} {Outcomes}
  Under {Prospective} {Payment},''} \emph{Econometrica} 63, no. 1
  (1995): 29.
\item
  Randall P Ellis and Thomas G McGuire {``Hospital Response to
  Prospective Payment: Moral Hazard, Selection, and Practice-Style
  Effects,''} \emph{Journal of Health Economics} 15, no. 3 (1996):
  257--277.
\item
  Elaine Silverman and Jonathan Skinner {``Medicare Upcoding and
  Hospital Ownership,''} \emph{Journal of Health Economics} 23, no. 2
  (2004): 369--389.
\item
  Caitlin Carroll et al. {``Effects of Episode-Based Payment on Health
  Care Spending and Utilization: {Evidence} from Perinatal Care in
  {Arkansas},''} \emph{Journal of Health Economics} 61 (2018): 47--62.
\item
  Amanda Cook and Susan Averett {``Do {Hospitals} {Respond} to
  {Changing} {Incentive} {Structures}? {Evidence} from {Medicare}'s 2007
  {DRG} {Restructuring},''} \emph{Journal of Health Economics} (2020):
  102319.
\item
  Sunita Desai and J Michael McWilliams {``Consequences of the {340b}
  Drug Pricing Program,''} \emph{New England Journal of Medicine} 378,
  no. 6 (2018): 539--548.
\item
  Gautam Gowrisankaran, Keith A Joiner, and Jianjing Lin \emph{How Do
  {Hospitals} {Respond} to {Payment} {Incentives}?} (National Bureau of
  Economic Research, 2019).
\item
  Atul Gupta {``Impacts of {Performance} {Pay} for {Hospitals}: {The}
  {Readmissions} {Reduction} {Program},''} \emph{American Economic
  Review} 111, no. 4 (April 2021): 1241--1283.
\end{itemize}

\hypertarget{class-04-0905-labor-day-no-class}{%
\subsubsection{Class 04, 09/05: Labor Day, no
class}\label{class-04-0905-labor-day-no-class}}

\hypertarget{class-05-0907-hospitals-and-financial-incentives-part-ii}{%
\subsubsection{Class 05, 09/07: Hospitals and financial incentives, part
II}\label{class-05-0907-hospitals-and-financial-incentives-part-ii}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Batty and Ippolito {``Financial Incentives, Hospital Care, and Health
  Outcomes.''}
\item
  Lee {``How Do Hospitals Respond to Managed Care? {Evidence} from
  at-Risk Newborns.''}
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  See supplemental reading in hospitals and financial incentives, part I
\end{itemize}

\hypertarget{class-06-0912---class-08-0919-application}{%
\subsubsection{Class 06, 09/12 - Class 08, 09/19:
Application}\label{class-06-0912---class-08-0919-application}}

\begin{itemize}
\tightlist
\item
  \textbf{Question:} How does insurance expansion affect hospital
  uncompensated care?
\item
  \textbf{Research design:} Difference-in-differences
\end{itemize}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Brantly Callaway and Pedro H. C. Sant'Anna
  {``Difference-in-{Differences} with Multiple Time Periods,''}
  \emph{Journal of Econometrics} 225, no. 2, Themed {Issue}: {Treatment}
  {Effect} 1 (December 2021): 200--230.
\item
  Liyang Sun and Sarah Abraham {``Estimating Dynamic Treatment Effects
  in Event Studies with Heterogeneous Treatment Effects,''}
  \emph{Journal of Econometrics} 225, no. 2, Themed {Issue}: {Treatment}
  {Effect} 1 (December 2021): 175--199.
\item
  Ashesh Rambachan and Jonathan Roth {``A {More} {Credible} {Approach}
  to {Parallel} {Trends},''} \emph{Working Paper} (2022).
\item
  Jonathan Roth et al. {``What's {Trending} in
  {Difference}-in-{Differences}? {A} {Synthesis} of the {Recent}
  {Econometrics} {Literature},''} \emph{arXiv Preprint arXiv:2201.01194}
  (2022).
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Too many to list! If you want to get up to speed on DD, please take a
  look at this
  \href{https://asjadnaqvi.github.io/DiD/docs/reading/04_literature/}{DiD
  Project}. It has code, lit review, and links to other videos and blog
  posts.
\item
  Dmitry Arkhangelsky et al. {``Synthetic
  {Difference}-in-{Differences},''} \emph{American Economic Review} 111,
  no. 12 (December 2021): 4088--4118.
\end{itemize}

\hypertarget{module-2-physicians}{%
\subsection{Module 2: Physicians}\label{module-2-physicians}}

\hypertarget{class-09-0921-background-and-institutional-details}{%
\subsubsection{Class 09, 09/21: Background and institutional
details**}\label{class-09-0921-background-and-institutional-details}}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Thomas G McGuire {``Physician Agency,''} \emph{Handbook of Health
  Economics} 1 (2000): 461--536.
\item
  Sanford J Grossman and Oliver D Hart {``An Analysis of the
  Principal-Agent Problem,''} \emph{Econometrica} (1983): 7--45.
\item
  John Wennberg and Alan Gittelsohn {``Small Area Variations in Health
  Care Delivery: A Population-Based Health Information System Can Guide
  Planning and Regulatory Decision-Making,''} \emph{Science} 182, no.
  4117 (1973): 1102--1108.
\item
  John E. Wennberg, Elliott S. Fisher, and Jonathan S. Skinner
  {``Geography {And} {The} {Debate} {Over} {Medicare} {Reform},''}
  \emph{Health Affairs} (2004): W96--W114.
\item
  Jonathan Skinner and Elliott Fisher {``Regional Disparities in
  {Medicare} Expenditures: An Opportunity for Reform,''} \emph{National
  Tax Journal} 50, no. 3 (1997): 413--425.
\item
  Amy Finkelstein, Matthew Gentzkow, and Heidi Williams {``Sources of
  Geographic Variation in Health Care: {Evidence} from Patient
  Migration,''} \emph{The Quarterly Journal of Economics} 131, no. 4
  (2016): 1681--1726.
\item
  David Molitor {``The Evolution of Physician Practice Styles: Evidence
  from Cardiologist Migration,''} \emph{American Economic Journal:
  Economic Policy} 10, no. 1 (2018): 326--56.
\end{itemize}

\hypertarget{class-10-0926-physician-agency-financial-incentives-and-referrals}{%
\subsubsection{Class 10, 09/26: Physician agency, financial incentives,
and
referrals}\label{class-10-0926-physician-agency-financial-incentives-and-referrals}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Jeffrey Clemens and Joshua D Gottlieb {``Do {Physicians}' {Financial}
  {Incentives} {Affect} {Medical} {Treatment} and {Patient} {Health}?''}
  \emph{American Economic Review} 104, no. 4 (2014): 1320--1349.
\item
  Dan Zeltzer {``Gender {Homophily} in {Referral} {Networks}:
  {Consequences} for the {Medicare} {Physician} {Earnings} {Gap},''}
  \emph{American Economic Journal: Applied Economics} 12, no. 2 (April
  2020): 169--97.
\end{itemize}

\textbf{Supplemental reading::}

\begin{itemize}
\tightlist
\item
  Laurence C Baker, M Kate Bundorf, and Daniel P Kessler {``The {Effect}
  of {Hospital}/{Physician} {Integration} on {Hospital} {Choice},''}
  \emph{Journal of Health Economics} 50 (2016): 1--8.
\item
  Jonathan Gruber and Maria Owings {``Physician Financial Incentives and
  Cesarean Section Delivery,''} \emph{The RAND Journal of Economics} 27,
  no. 1 (1996): 99--123.
\item
  Christopher C. Afendulis and Daniel P. Kessler {``Tradeoffs from
  {Integrating} {Diagnosis} and {Treatment} in {Markets} for {Health}
  {Care},''} \emph{American Economic Review} 97, no. 3 (June 2007):
  1013--1020.
\item
  Toshiaki Iizuka {``Physician Agency and Adoption of Generic
  Pharmaceuticals,''} \emph{American Economic Review} 102, no. 6 (2012):
  2826--58.
\item
  Christopher M Whaley et al. {``Higher {Medicare} {Spending} on
  {Imaging} and {Lab} {Services} After {Primary} {Care} {Physician}
  {Group} {Vertical} {Integration},''} \emph{Health Affairs} 40, no. 5
  (2021): 702--709.
\item
  Gary J Young et al. {``Hospital {Employment} of {Physicians} in
  {Massachusetts} {Is} {Associated} with {Inappropriate} {Diagnostic}
  {Imaging},''} \emph{Health Affairs} 40, no. 5 (2021): 710--718.
\item
  Emily Walden \emph{Can {Hospitals} {Buy} {Referrals}? {The} {Impact}
  of {Physician} {Group} {Acquisitions} on {Market}-{Wide} {Referral}
  {Patterns}}, Working \{Paper\} (University of Wisconsin, Madison,
  2016).
\item
  Michael R. Richards, Jonathan A. Seward, and Christopher M. Whaley
  {``Treatment Consolidation After Vertical Integration: {Evidence} from
  Outpatient Procedure Markets,''} \emph{Journal of Health Economics} 81
  (January 2022): 102569.
\item
  Baker, Bundorf, and Kessler {``The {Effect} of {Hospital}/{Physician}
  {Integration} on {Hospital} {Choice}.''}
\item
  Michael Chernew et al. {``Physician Agency, Consumerism, and the
  Consumption of Lower-Limb {MRI} Scans,''} \emph{Journal of Health
  Economics} 76 (March 2021): 102427.
\item
  Thomas G McGuire and Mark V Pauly {``Physician Response to Fee Changes
  with Multiple Payers,''} \emph{Journal of Health Economics} 10, no. 4
  (January 1991): 385--410.
\item
  Kate Ho and Ariel Pakes {``Hospital {Choices}, {Hospital} {Prices},
  and {Financial} {Incentives} to {Physicians},''} \emph{The American
  Economic Review} 104, no. 12 (December 2014): 3841--3884.
\end{itemize}

\hypertarget{class-11-0928-physician-learning}{%
\subsection{Class 11, 09/28: Physician
learning}\label{class-11-0928-physician-learning}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Gregory S Crawford and Matthew Shum {``Uncertainty and Learning in
  Pharmaceutical Demand,''} \emph{Econometrica} 73, no. 4 (2005):
  1137--1173.
\item
  Heather Sarsons \emph{Interpreting {Signals} in the {Labor} {Market}:
  {Evidence} from {Medical} {Referrals}}, Working \{Paper\} (Harvard
  University, 2018).
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Andrea Coscelli and Matthew Shum {``An Empirical Model of Learning and
  Patient Spillovers in New Drug Entry,''} \emph{Journal of
  Econometrics} 122, no. 2 (2004): 213--246.
\item
  Maria Marta Ferreyra and Grigory Kosenok {``Learning about {New}
  {Products}: {An} {Empirical} {Study} of {Physician}'s {Behavior},''}
  \emph{Economic Inquiry} 49, no. 3 (2011): 876--898.
\item
  Tat Chan, Chakravarthi Narasimhan, and Ying Xie {``Treatment
  Effectiveness and Side Effects: {A} Model of Physician Learning,''}
  \emph{Management Science} 59, no. 6 (2013): 1309--1325.
\item
  Michael Dickstein \emph{Efficient Provision of Experience Goods:
  {Evidence} from Antidepressant Choice}, Working \{Paper\} (New York
  University, Leonard N. Stern School of Business, 2018).
\item
  Qing Gong \emph{Physician {Learning} and {Treatment} {Choices}
  {Evidence} from {Brain} {Aneurysms}}, Working \{Paper\} (University of
  North Carolina at Chapel Hill, 2018).
\item
  Erin M Johnson \emph{Ability, {Learning} and the {Career} {Path} of
  {Cardiac} {Specialists}}, Working \{Paper\} (MIT, 2011).
\item
  Vivian Ho {``Learning and the Evolution of Medical Technologies: The
  Diffusion of Coronary Angioplasty,''} \emph{Journal of Health
  Economics} 21, no. 5 (2002): 873--885.
\end{itemize}

\hypertarget{class-12-1003---class-13-1005-application}{%
\subsubsection{Class 12, 10/03 - Class 13, 10/05:
Application}\label{class-12-1003---class-13-1005-application}}

\begin{itemize}
\tightlist
\item
  \textbf{Question:} Physician effort and integration with hospitals
\item
  \textbf{Research design:} Instrumental variables
\end{itemize}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  David Dranove and Christopher Ody {``Employed for Higher Pay? {How}
  {Medicare} Facility Fees Affect Hospital Employment of Physicians,''}
  \emph{American Economics Journal: Economic Policy} 11, no. 4 (2019):
  249--271.
\item
  Kirill Borusyak and Peter Hull \emph{Non-Random Exposure to Exogenous
  Shocks: {Theory} and Applications} (National Bureau of Economic
  Research, 2021).
\item
  Emily Oster {``Unobservable {Selection} and {Coefficient} {Stability}:
  {Theory} and {Evidence},''} \emph{Journal of Business \& Economic
  Statistics} 37, no. 2 (2019): 187--204.
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Guido W. Imbens {``Better {LATE} {Than} {Nothing}: {Some} {Comments}
  on {Deaton} (2009) and {Heckman} and {Urzua} (2009),''} \emph{Journal
  of Economic Literature} 48, no. 2 (June 2010): 399--423.
\item
  Victor Chernozhukov et al. {``Double/Debiased Machine Learning for
  Treatment and Structural Parameters,''} \emph{The Econometrics
  Journal} 21, no. 1 (February 2018): C1--C68.
\item
  Joshua Angrist and Michal Kolesár {``One {Instrument} to {Rule} {Them}
  {All}: {The} {Bias} and {Coverage} of {Just}-{ID} {IV},''} Working
  \{Paper\}, Working {Paper} {Series} (National Bureau of Economic
  Research, October 2021).
\item
  James Heckman, Sergio Urzua, and Edward Vytlacil {``Understanding
  {Instrumental} {Variables} in {Models} with {Essential}
  {Heterogeneity},''} \emph{The Review of Economics and Statistics} 88
  (2006): 389--432.
\item
  Isaiah Andrews, James H Stock, and Liyang Sun {``Weak Instruments in
  Instrumental Variables Regression: {Theory} and Practice,''}
  \emph{Annual Review of Economics} 11 (2019): 727--753.
\item
  David S. Lee et al. {``Valid t-Ratio {Inference} for {IV},''} Working
  \{Paper\}, Working {Paper} {Series} (National Bureau of Economic
  Research, August 2021).
\item
  Kirill Borusyak, Peter Hull, and Xavier Jaravel
  {``Quasi-{Experimental} {Shift}-{Share} {Research} {Designs},''}
  \emph{Review of Economic Studies} (2021).
\end{itemize}

\hypertarget{class-14-1010-fall-break}{%
\subsubsection{Class 14, 10/10: Fall
break}\label{class-14-1010-fall-break}}

\hypertarget{class-15-1012-application-contd}{%
\subsubsection{Class 15, 10/12: Application,
cont'd}\label{class-15-1012-application-contd}}

\hypertarget{module-3-information-disclosure}{%
\subsection{Module 3: Information
disclosure}\label{module-3-information-disclosure}}

\hypertarget{class-16-1017-patient-decision-making}{%
\subsubsection{Class 16, 10/17: Patient
decision-making}\label{class-16-1017-patient-decision-making}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Jonathan D Ketcham et al. {``Sinking, Swimming, or Learning to Swim in
  {Medicare} {Part} {D},''} \emph{American Economic Review} 102, no. 6
  (2012): 2639--2673.
\item
  Martin Gaynor, Carol Propper, and Stephan Seiler {``Free to {Choose}?
  {Reform}, {Choice}, and {Consideration} {Sets} in the {English}
  {National} {Health} {Service},''} \emph{American Economic Review} 106,
  no. 11 (November 2016): 3521--57.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Joseph Farrell and Paul Klemperer {``Coordination and Lock-in:
  {Competition} with Switching Costs and Network Effects,''}
  \emph{Handbook of Industrial Organization} 3 (2007): 1967--2072.
\item
  Jason Abaluck and Jonathan Gruber {``Choice {Inconsistencies} Among
  the {Elderly}: {Evidence} from {Plan} {Choice} in the {Medicare}
  {Part} {D} {Program},''} \emph{American Economic Review} 101, no. 4
  (June 2011): 1180--1210.
\end{itemize}

\hypertarget{class-17-1019-inertia-in-health-insurance-choice}{%
\subsubsection{Class 17, 10/19: Inertia in health insurance
choice}\label{class-17-1019-inertia-in-health-insurance-choice}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Benjamin R. Handel {``Adverse {Selection} and {Inertia} in {Health}
  {Insurance} {Markets}: {When} {Nudging} {Hurts},''} \emph{American
  Economic Review} 103, no. 7 (December 2013): 2643--2682.
\item
  Keith M Ericson {``Consumer Inertia and Firm Pricing in the {Medicare}
  {Part} {D} Prescription Drug Insurance Exchange,''} \emph{American
  Economic Journal: Economic Policy} 6, no. 1 (2014): 38--64.
\end{itemize}

\hypertarget{class-18-1024-effects-of-quality-disclosure}{%
\subsubsection{Class 18, 10/24: Effects of quality
disclosure}\label{class-18-1024-effects-of-quality-disclosure}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  David Dranove et al. {``Is {More} {Information} {Better}? {The}
  {Effects} of {Report} {Cards} on {Health} {Care} {Providers},''}
  \emph{Journal of Political Economy} 111, no. 3 (2003): 555--588.
\item
  L. Dafny and D. Dranove {``Do Report Cards Tell Consumers Anything
  They Don't Already Know? {The} Case of {Medicare} {HMOs},''}
  \emph{RAND Journal of Economics} 39, no. 3 (2008): 790--821.
\item
  Jonathan T Kolstad {``Information and Quality When Motivation Is
  Intrinsic: {Evidence} from Surgeon Report Cards,''} \emph{American
  Economic Review} 103, no. 7 (2013): 2875--2910.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  M. Chernew, D. P. Scanlon, et al. {``Health Plan Report Cards and
  Insurance Choice.''} \emph{Inquiry} 35, no. 1 (1998): 9.
\item
  D. P. Scanlon et al. {``The Impact of Health Plan Report Cards on
  Managed Care Enrollment,''} \emph{Journal of Health Economics} 21, no.
  1 (2002): 19--41.
\item
  G. J. Wedig and M. Tai-Seale {``The Effect of Report Cards on Consumer
  Choice in the Health Insurance Market,''} \emph{Journal of Health
  Economics} 21, no. 6 (2002): 1031--1048.
\item
  G. Z. Jin and A. T. Sorensen {``Information and Consumer Choice: The
  Value of Publicized Health Plan Ratings,''} \emph{Journal of Health
  Economics} 25, no. 2 (2006): 248--275.
\item
  David Dranove and Ginger Zhe Jin {``Quality {Disclosure} and
  {Certification}: {Theory} and {Practice},''} \emph{Journal of Economic
  Literature} 48, no. 4 (2010): 935--963.
\item
  Rachel O Reid et al. {``Association {Between} {Medicare} {Advantage}
  {Plan} {Star} {Ratings} and {EnrollmentStar} {Ratings} for {Medicare}
  {Advantage} {Plan},''} \emph{JAMA} 309, no. 3 (2013): 267--274.
\item
  M. Darden and I. McCarthy {``The {Star} {Treatment}: {Estimating} the
  {Impact} of {Star} {Ratings} on {Medicare} {Advantage}
  {Enrollments},''} \emph{Journal of Human Resources} 50, no. 4 (2015):
  980--1008.
\item
  I. McCarthy and M. Darden {``Supply-Side {Responses} to {Public}
  {Quality} {Ratings}: {Evidence} from {Medicare} {Advantage},''}
  \emph{American Journal of Health Economics} 3, no. 2 (2017): 140--164.
\item
  Andrew J. Epstein {``Effects of Report Cards on Referral Patterns to
  Cardiac Surgeons,''} \emph{Journal of Health Economics} 29, no. 5
  (September 2010): 718--731.
\end{itemize}

\hypertarget{class-19-1026-effects-of-price-disclosure}{%
\subsubsection{Class 19, 10/26: Effects of price
disclosure}\label{class-19-1026-effects-of-price-disclosure}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Zach Y. Brown {``Equilibrium {Effects} of {Health} {Care} {Price}
  {Information},''} \emph{The Review of Economics and Statistics} 101,
  no. 4 (October 2019): 699--712.
\item
  Matthew Grennan and Ashley Swanson {``Transparency and Negotiated
  Prices: {The} Value of Information in Hospital-Supplier Bargaining,''}
  \emph{Journal of Political Economy} 128, no. 4 (2020): 1234--1268.
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Chernew et al. {``Physician Agency, Consumerism, and the Consumption
  of Lower-Limb {MRI} Scans.''}
\item
  Zarek C. Brot-Goldberg et al. {``What Does a {Deductible} {Do}? {The}
  {Impact} of {Cost}-{Sharing} on {Health} {Care} {Prices},
  {Quantities}, and {Spending} {Dynamics}*,''} \emph{The Quarterly
  Journal of Economics} 132, no. 3 (August 2017): 1261--1318.
\end{itemize}

\hypertarget{class-20-1031-and-class-21-1102-application}{%
\subsubsection{Class 20, 10/31 and Class 21, 11/02:
Application}\label{class-20-1031-and-class-21-1102-application}}

\begin{itemize}
\tightlist
\item
  \textbf{Question:} Consumer inertia and insurer premiums
\item
  \textbf{Research design:} Regression discontinuity
\end{itemize}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Ericson {``Consumer Inertia and Firm Pricing in the {Medicare} {Part}
  {D} Prescription Drug Insurance Exchange.''}
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Best starting place for most up-to-date info on RD is from Matias
  Cattaneo and team. Their site, \href{https://rdpackages.github.io/}{RD
  Packages}, has a tone of great info with links to the most recent
  literature.
\end{itemize}

\hypertarget{module-4-healthcare-competition-and-industrial-organization}{%
\subsection{Module 4: Healthcare competition and industrial
organization}\label{module-4-healthcare-competition-and-industrial-organization}}

\hypertarget{class-22-1107-competition-in-price-and-quality}{%
\subsubsection{Class 22, 11/07: Competition in price and
quality**}\label{class-22-1107-competition-in-price-and-quality}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  David Dranove, Mark Shanley, and Carol Simon {``Is Hospital
  Competition Wasteful?''} \emph{RAND Journal of Economics} (1992):
  247--262.
\item
  Ying Fan {``Ownership Consolidation and Product Characteristics: {A}
  Study of the {US} Daily Newspaper Market,''} \emph{American Economic
  Review} 103, no. 5 (2013): 1598--1628.
\item
  Matthew Lewis and Kevin Pflum {``Competition and {Quality} {Choice} in
  {Hospital} {Markets},''} \emph{Working Paper} (2016).
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  A Spence {``Monopoly, {Quality}, and {Regulation},''} \emph{Bell
  Journal of Economics} 6, no. 2 (1975): 417--429.
\item
  Michael Mussa and Sherwin Rosen {``Monopoly and Product Quality,''}
  \emph{Journal of Economic Theory} 18, no. 2 (1978): 301--317.
\item
  Jerry Hausman, Gregory Leonard, and J Douglas Zona {``Competitive
  Analysis with Differentiated Products,''} \emph{Annales d'Economie Et
  de Statistique} (1994): 159--180.
\item
  James D Dana Jr and Yuk-Fai Fong {``Product Quality, Reputation, and
  Market Structure,''} \emph{International Economic Review} 52, no. 4
  (2011): 1059--1076.
\end{itemize}

\hypertarget{class-23-1109-measuring-hospital-markets-and-market-power}{%
\subsubsection{Class 23, 11/09: Measuring hospital markets and market
power}\label{class-23-1109-measuring-hospital-markets-and-market-power}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  David Dranove and Christopher Ody {``Evolving {Measures} of {Provider}
  {Market} {Power},''} \emph{American Journal of Health Economics} 2,
  no. 2 (2016): 145--160.
\item
  Jordan Everson, John M Hollingsworth, and Julia Adler-Milstein
  {``Comparing Methods of Grouping Hospitals,''} \emph{Health Services
  Research} 54, no. 5 (2019): 1090--1098.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  David Dranove and Mark Shanley {``A Note on the Relational Aspects of
  Hospital Market Definitions,''} \emph{Journal of Health Economics} 8,
  no. 4 (1990): 473--478.
\item
  Cory Capps, David Dranove, and Mark Satterthwaite {``Competition and
  Market Power in Option Demand Markets,''} \emph{RAND Journal of
  Economics} (2003): 737--763.
\end{itemize}

\hypertarget{class-24-1114-structure-conduct-performance}{%
\subsubsection{Class 24, 11/14:
``Structure-Conduct-Performance''}\label{class-24-1114-structure-conduct-performance}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Daniel Kessler and Mark McClellan {``Is Hospital Competition Socially
  Wasteful?''} \emph{Quarterly Journal of Economics} 2, no. 115 (2000):
  577--615.
\item
  Nathan Miller et al. {``On the Misuse of Regressions of Price on the
  {HHI} in Merger Review,''} \emph{Journal of Antitrust Enforcement} 10,
  no. 2 (July 2022): 248--259.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Dranove, Shanley, and Simon {``Is Hospital Competition Wasteful?''}
\item
  William J Lynk {``Nonprofit Hospital Mergers and the Exercise of
  Market Power,''} \emph{The Journal of Law and Economics} 38, no. 2
  (1995): 437--461.
\item
  Emmett B Keeler, Glenn Melnick, and Jack Zwanziger {``The Changing
  Effects of Competition on Non-Profit and for-Profit Hospital Pricing
  Behavior,''} \emph{Journal of Health Economics} 18, no. 1 (1999):
  69--86.
\item
  Zack Cooper et al. {``The Price Ain't Right? {Hospital} Prices and
  Health Spending on the Privately Insured,''} \emph{Quarterly Journal
  of Economics} 134, no. 1 (2019): 51--107.
\end{itemize}

\hypertarget{class-25-1116-bargaining-framework}{%
\subsubsection{Class 25, 11/16: Bargaining
framework}\label{class-25-1116-bargaining-framework}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Kate Ho and Robin S Lee {``Insurer Competition in Health Care
  Markets,''} \emph{Econometrica} 85, no. 2 (2017): 379--417.
\item
  Jeffrey Clemens and Joshua D Gottlieb {``In the Shadow of a Giant:
  {Medicare}'s Influence on Private Physician Payments,''} \emph{Journal
  of Political Economy} 125, no. 1 (2017): 1--39.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Timothy F Bresnahan and Peter C Reiss {``Entry and Competition in
  Concentrated Markets,''} \emph{Journal of Political Economy} (1991):
  977--1009.
\item
  Peter C. Reiss and F. Wolak {``Structural Econometric Modeling:
  {Rationales} and Examples from Industrial Organization. {Handbook} of
  {Econometrics},''} \emph{Handbook of Econometrics, Edition} 1 (2007).
\item
  Jean Abraham, Martin Gaynor, and William B Vogt {``Entry and
  {Competition} in {Local} {Hospital} {Markets},''} \emph{Journal of
  Industrial Economics} 55, no. 2 (2007): 265--288.
\item
  Gautam Gowrisankaran, Aviv Nevo, and Robert Town {``Mergers {When}
  {Prices} {Are} {Negotiated}: {Evidence} from the {Hospital}
  {Industry},''} \emph{American Economic Review} 105, no. 1 (2015):
  172--203.
\item
  Matthew Lewis and Kevin Pflum {``Diagnosing {Hospital} {System}
  {Bargaining} {Power} in {Managed} {Care} {Networks},''} \emph{American
  Economic Journal: Economic Policy} 7, no. 1 (2015): 243--274.
\end{itemize}

\hypertarget{class-26-1121-no-meeting-today.-class-rescheduled-for-december-2}{%
\subsubsection{Class 26, 11/21: No meeting today. Class rescheduled for
December
2}\label{class-26-1121-no-meeting-today.-class-rescheduled-for-december-2}}

\hypertarget{class-27-1123-effects-of-mergersclosures-virtual-class}{%
\subsubsection{\texorpdfstring{Class 27, 11/23: Effects of
mergers/closures (\textbf{virtual
class})}{Class 27, 11/23: Effects of mergers/closures (virtual class)}}\label{class-27-1123-effects-of-mergersclosures-virtual-class}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Stuart V. Craig, Matthew Grennan, and Ashley Swanson {``Mergers and
  Marginal Costs: {New} Evidence on Hospital Buyer Power,''} \emph{The
  RAND Journal of Economics} 52, no. 1 (2021): 151--178.
\item
  Paul J Eliason et al. {``How {Acquisitions} {Affect} {Firm} {Behavior}
  and {Performance}: {Evidence} from the {Dialysis} {Industry},''}
  \emph{Quarterly Journal of Economics} 135, no. 1 (2020): 221--267.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Michael G Vita and Seth Sacher {``The Competitive Effects of
  Not-for-Profit Hospital Mergers: A Case Study,''} \emph{Journal of
  Industrial Economics} 49, no. 1 (2001): 63--84.
\item
  Gaynor and Vogt {``Competition Among {Hospitals}.''}
\item
  Richard C Lindrooth, Anthony T Lo Sasso, and Gloria J Bazzoli {``The
  Effect of Urban Hospital Closure on Markets,''} \emph{Journal of
  Health Economics} 22, no. 5 (2003): 691--712.
\item
  M Gaynor and R Town \emph{The Impact of Hospital Consolidation -
  {Update}}, Policy \{Brief\} (Robert Wood Johnson Foundation, 2012).
\item
  Leemore Dafny {``Estimation and {Identification} of {Merger}
  {Effects}: {An} {Application} to {Hospital} {Mergers},''}
  \emph{Journal of Law and Economics} 52, no. 3 (2009): 523--550.
\item
  Martin Gaynor, Rodrigo Moreno-Serra, and Carol Propper {``Death by
  Market Power: Reform, Competition, and Patient Outcomes in the
  {National} {Health} {Service},''} \emph{American Economic Journal:
  Economic Policy} 5, no. 4 (2013): 134--66.
\end{itemize}

\hypertarget{class-28-1128-out-of-market-mergers}{%
\subsubsection{Class 28, 11/28: Out-of-market
mergers}\label{class-28-1128-out-of-market-mergers}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Leemore Dafny, Kate Ho, and Robin S Lee {``The Price Effects of
  Cross-Market Mergers: Theory and Evidence from the Hospital
  Industry,''} \emph{RAND Journal of Economics} 50, no. 2 (2019):
  286--325.
\item
  Matt Schmitt {``Multimarket {Contact} in the {Hospital} {Industry},''}
  \emph{American Economic Journal: Economic Policy} 10, no. 3 (2018):
  361--87.
\end{itemize}

\hypertarget{class-29-1130-vertical-integration}{%
\subsubsection{Class 29, 11/30: Vertical
integration}\label{class-29-1130-vertical-integration}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Fernando Luco and Guillermo Marshall {``The {Competitive} {Impact} of
  {Vertical} {Integration} by {Multiproduct} {Firms},''} \emph{American
  Economic Review} 110, no. 7 (July 2020): 2041--2064.
\item
  Thomas G. Koch, Brett W. Wendling, and Nathan E. Wilson {``The
  {Effects} of {Physician} and {Hospital} {Integration} on {Medicare}
  {Beneficiaries}' {Health} {Outcomes},''} \emph{The Review of Economics
  and Statistics} 103, no. 4 (September 2021): 725--739.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Dranove and Ody {``Employed for Higher Pay? {How} {Medicare} Facility
  Fees Affect Hospital Employment of Physicians.''}
\item
  Thomas G Koch, Brett W Wendling, and Nathan E Wilson {``How Vertical
  Integration Affects the Quantity and Cost of Care for {Medicare}
  Beneficiaries,''} \emph{Journal of Health Economics} 52 (2017):
  19--32.
\item
  Haizhen Lin, Ian M. McCarthy, and Michael Richards {``Hospital
  {Pricing} {Following} {Integration} with {Physician} {Practices},''}
  \emph{Journal of Health Economics} 77 (May 2021): 102444.
\item
  Lawton R Burns and Mark V Pauly {``Integrated Delivery Networks: A
  Detour on the Road to Integrated Health Care?''} \emph{Health Affairs}
  21, no. 4 (2002): 128--143.
\item
  Federico Ciliberto and David Dranove {``The Effect of
  Physician--Hospital Affiliations on Hospital Prices in
  {California},''} \emph{Journal of Health Economics} 25, no. 1 (2006):
  29--38.
\item
  Alison Evans Cuellar and Paul J Gertler {``Strategic Integration of
  Hospitals and Physicians,''} \emph{Journal of Health Economics} 25,
  no. 1 (2006): 1--28.
\item
  Baker, Bundorf, and Kessler {``The {Effect} of {Hospital}/{Physician}
  {Integration} on {Hospital} {Choice}.''}
\item
  Hannah T Neprash et al. {``Association of Financial Integration
  Between Physicians and Hospitals with Commercial Health Care
  Prices,''} \emph{JAMA Internal Medicine} 175, no. 12 (2015):
  1932--1939.
\item
  Cory Capps, David Dranove, and Christopher Ody {``The Effect of
  Hospital Acquisitions of Physician Practices on Prices and
  Spending,''} \emph{Journal of Health Economics} 59 (2018): 139--152.
\item
  R Tamara Konetzka, Elizabeth A Stuart, and Rachel M Werner {``The
  Effect of Integration of Hospitals and Post-Acute Care Providers on
  {Medicare} Payment and Patient Outcomes,''} \emph{Journal of Health
  Economics} (2018).
\end{itemize}

\hypertarget{rescheduled-from-1121-and-class-30-1205-application}{%
\subsubsection{12/02 (rescheduled from 11/21) and Class 30, 12/05:
Application}\label{rescheduled-from-1121-and-class-30-1205-application}}

\begin{itemize}
\tightlist
\item
  \textbf{Question:} Sensitivity of demand estimates to market
  definition
\item
  \textbf{Research design:} Structural demand estimation
\end{itemize}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Everson, Hollingsworth, and Adler-Milstein {``Comparing Methods of
  Grouping Hospitals.''}
\end{itemize}

\textbf{Supplemental reading:}

\begin{itemize}
\tightlist
\item
  Aviv Nevo {``A Practitioner's Guide to Estimation of
  Random-Coefficients Logit Models of Demand,''} \emph{Journal of
  Economics \& Management Strategy} 9, no. 4 (2000): 513--548.
\item
  Christopher Conlon and Jeff Gortmaker {``Best Practices for
  Differentiated Products Demand Estimation with {PyBLP},''} \emph{The
  RAND Journal of Economics} 51, no. 4 (2020): 1108--1161.
\end{itemize}

\hypertarget{module-5-health-insurance-markets-we-wont-get-to-this}{%
\subsection{Module 5: Health insurance markets (we won't get to
this)}\label{module-5-health-insurance-markets-we-wont-get-to-this}}

\hypertarget{insurance-markets-and-competitiveness}{%
\subsubsection{Insurance markets and
competitiveness}\label{insurance-markets-and-competitiveness}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Leemore Dafny {``Are {Health} {Insurance} {Markets} {Competitive}?''}
  \emph{American Economic Review} 100, no. 4 (2010): 1399--1431.
\item
  Leemore Dafny, Mark Duggan, and Subramaniam Ramanarayanan {``Paying a
  {Premium} on {Your} {Premium}? {Consolidation} in the {US} {Health}
  {Insurance} {Industry},''} \emph{American Economic Review} 102, no. 2
  (2012): 1161--1185.
\item
  M Kate Bundorf, Jonathan Levin, and Neale Mahoney {``Pricing and
  Welfare in Health Plan Choice,''} \emph{American Economic Review} 102,
  no. 7 (2012): 3214--3248.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Robert Town and Su Liu {``The Welfare Impact of {Medicare} {HMOs},''}
  \emph{RAND Journal of Economics} 24, no. 4 (2003): 719--736.
\item
  Liran Einav, Amy Finkelstein, and Jonathan Levin {``Beyond {Testing}:
  {Empirical} {Models} of {Insurance} {Markets},''} \emph{Annual Review
  of Economics} 2, no. 1 (2010): 311--336.
\item
  Liran Einav, Amy Finkelstein, and Mark R Cullen {``Estimating
  {Welfare} in {Insurance} {Markets} {Using} {Variation} in {Prices},''}
  \emph{Quarterly Journal of Economics} 125, no. 3 (2010): 877--921.
\item
  Amanda Starc {``Insurer Pricing and Consumer Welfare: {Evidence} from
  Medigap,''} \emph{RAND Journal of Economics} 45, no. 1 (2014):
  198--220.
\end{itemize}

\hypertarget{adverse-selection}{%
\subsubsection{Adverse selection}\label{adverse-selection}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  David M Cutler and Sarah J Reber {``Paying for {Health} {Insurance}:
  {The} {Trade}-{Off} Between {Competition} and {Adverse}
  {Selection},''} \emph{Quarterly Journal of Economics} 113, no. 2
  (1998): 433--466.
\item
  Francesco Decarolis and Andrea Guglielmo {``Insurers' Response to
  Selection Risk: {Evidence} from {Medicare} Enrollment Reforms,''}
  \emph{Journal of Health Economics} 56 (2017): 383--396.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  George Ackerloff {``The Market for Lemons: {Quality} Uncertainty and
  the Market Mechanism,''} \emph{Quarterly Journal of Economics} 84, no.
  3 (1970): 488--500.
\item
  Michael Rothschild and Joseph Stiglitz {``Equilibrium in Competitive
  Insurance Markets: {An} Essay on the Economics of Imperfect
  Information,''} \emph{Quarterly Journal of Economics} (1976):
  629--649.
\item
  Richard G Frank, Jacob Glazer, and Thomas G McGuire {``Measuring
  Adverse Selection in Managed Health Care,''} \emph{Journal of Health
  Economics} 19, no. 6 (2000): 829--854.
\item
  Liran Einav and Amy Finkelstein {``Selection in {Insurance} {Markets}:
  {Theory} and {Empirics} in {Pictures},''} \emph{The Journal of
  Economic Perspectives} 25, no. 1 (2011): 115--138.
\end{itemize}

\hypertarget{moral-hazard}{%
\subsubsection{Moral hazard}\label{moral-hazard}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Liran Einav et al. {``Selection on Moral Hazard in Health
  Insurance,''} \emph{American Economic Review} 103, no. 1 (2013):
  178--219.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Willard G Manning and M Susan Marquis {``Health Insurance: The
  Tradeoff Between Risk Pooling and Moral Hazard,''} \emph{Journal of
  Health Economics} 15, no. 5 (1996): 609--639.
\item
  Amy Finkelstein \emph{Moral Hazard in Health Insurance} (Columbia
  University Press, 2014).
\item
  Liran Einav and Amy Finkelstein {``Moral {Hazard} in {Health}
  {Insurance}: {What} {We} {Know} and {How} {We} {Know} {It},''}
  \emph{Journal of the European Economic Association} 16, no. 4 (August
  2018): 957--982.
\end{itemize}

\hypertarget{managed-competition}{%
\subsubsection{Managed competition}\label{managed-competition}}

\textbf{Primary reading:}

\begin{itemize}
\tightlist
\item
  Vilsa Curto et al. {``Can {Health} {Insurance} {Competition} {Work}?
  {Evidence} from {Medicare} {Advantage},''} \emph{Journal of Political
  Economy} 129, no. 2 (2021): 570--606.
\item
  Liran Einav and Jonathan Levin {``Managed Competition in Health
  Insurance,''} \emph{Journal of the European Economic Association} 13,
  no. 6 (2015): 998--1021.
\end{itemize}

\textbf{Supplemental:}

\begin{itemize}
\tightlist
\item
  Zirui Song, Mary Beth Landrum, and Michael E Chernew {``Competitive
  Bidding in {Medicare} {Advantage}: {Effect} of Benchmark Changes on
  Plan Bids,''} \emph{Journal of Health Economics} 32, no. 6 (2013):
  1301--1312.
\item
  Marika Cabral, Michael Geruso, and Neale Mahoney {``Do Larger Health
  Insurance Subsidies Benefit Patients or Producers? {Evidence} from
  {Medicare} {Advantage},''} \emph{American Economic Review} 108, no. 8
  (2018): 2048--87.
\item
  Karen Stockley et al. \emph{Premium {Transparency} in the {Medicare}
  {Advantage} {Market}: {Implications} for {Premiums}, {Benefits}, and
  {Efficiency}}, Working \{Paper\} (National Bureau of Economic
  Research, 2014).
\item
  Mark Duggan, Amanda Starc, and Boris Vabson {``Who Benefits When the
  Government Pays More? {Pass}-Through in the {Medicare} {Advantage}
  Program,''} \emph{Journal of Public Economics} 141 (2016): 50--67.
\item
  Daria Pelech {``Paying More for Less? {Insurer} Competition and Health
  Plan Generosity in the {Medicare} {Advantage} Program,''}
  \emph{Journal of Health Economics} 61 (2018): 77--92.
\end{itemize}




\end{document}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother