---
title: "Module 4 Empirical Exercise"
author: "Econ 771: Health Economics II"
date: "Due: Tuesday, November 24"
output: 
  bookdown::html_document2:
    toc: TRUE
    toc_float: TRUE
    theme: darkly
    number_sections: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
In this assignment, we're going to work through some applied issues related to measuring hospital markets and estimating demand with market-level data.

To begin, please "accept" the assignment on our GitHub classroom page. This will initialize a GitHub repository where you can edit and ultimately craft your answers in your data management/econometrics software of choice.  

You will "turn in" your repository to me with your answers and all supporting documents. Please create a final document with your main answers and analyses in a PDF. Please also be sure to include in your repository all of your supporting code files. Practice writing good code and showing me only what I would need to recreate your results.


# Resources and data
The data for this assignment comes from three sources:

1. **HCRIS**: All hospitals that treat Medicare patients must submit a hospital cost report. The data are then made available on the CMS website. The NBER also provides a slightly more user-friendly version of these data. Info and code for these data are available at my GitHub page [here](https://github.com/imccart/HCRIS). The raw data are available on our AWS AMI in the "HCRIS" folder. Use these data for information on total discharges and to construct a measure of hospital prices.

2. **Hospital Market GitHub Repo**: John Graves has an excellent GitHub repository detailing the issues in measuring hospital markets. You can access that repo [here](https://github.com/graveja0/health-care-markets). To replicate some of his work using the community detection algorithm, you'll need to use the [Hospital Service Area Files](https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Hospital-Service-Area-File/index). The link takes you to the CMS website where you can download the data directly. I've also put these data in the "HSAF" folder on our AWS AMI. Note that much of John's code deals with presenting the markets on a map, rather than physically constructing the markets empirically. For a condensed version of the code focusing more on the market formation rather than mapping, see my GitHub repo [here](https://github.com/imccart/hospital-markets).


Note: Given the time crunch of the semester and the complexity of some of the raw data work for this assignment, I'm going to try to give you all a break here by skipping some steps for you. On our Canvas site, you'll find the 'hospital_markets.rds' data, which is the output from the hospital market github repo. I've also uploaded a couple of other geographic data files to give you info on HRR, zip, and fips. Once you've organized the HCRIS data following the repo listed in (1) above, ese the following code chunk to combine all of the data for this assignment:

```{r, eval=FALSE}
hcris.data <- readRDS(paste0(hcris.path,'/HCRIS_Data.rds'))
hrr.xw <- read_excel(path=paste0(geog.path,'/ZipHSAHrr15.xls'))
comm.det <- readRDS(paste0(community.path,'/hospital_markets.rds')) %>%
  mutate(fips=as.numeric(fips))
zip.county <- read_csv(paste0(geog.path,"/zcta-to-county.csv")) %>% 
  filter(row_number() != 1) %>% 
  mutate(zip = as.numeric(zcta5),
         fips = as.numeric(county),
         pct_zip_fips = as.numeric(afact)) %>% 
  as_tibble() %>%
  select(zip, fips, pct_zip_fips)
```


# Questions
In your GitHub repository, please be sure to clearly address/answer the following questions. In all cases, limit your data to the years 2000 through 2017.

1. Calculate hospital market shares when defining the "market" by zip code and using hospital discharges as your measure of quantity. Create a "violin plot" of market shares over time.

2. Calculate hospital market shares when defining the "market" as the HRR, again using hospital discharges as your measure of quantity. Create a violin plot of market shares over time. How do these results differ from part 1?

3. Calculate hospital market shares when defining the market based on the community detection algorithms described [here](https://github.com/graveja0/health-care-markets). Continue to use hospital discharges as your measure of quantity, and summarize your results with a violin plot. How do these results differ from those in parts 1 and 2?

4. Plot the average market shares over time for each of the three market definitions you've considered. Be sure to put all plots on the same graph.

5. Estimate a logit discrete choice model using market share data based on the different definition of markets in parts 1-3. Be sure to include year and hospital fixed effects in your estimation. What is your price elasticity estimate in each case, and how does it differ based on your measure of the market? *hint: don't worry about instrumenting for anything right now.*

